Detecting Spam on Social Networking Sites

### 2012-05-09 docs/paper/paper.tex
update

Unlike spam messages, we did not have access to a comparable corpus of human-labeled ham messages. To simplify the labeling of ham messages, we made the assumption that messages remaining in the inboxes of active users after a period of several months would most likely not be spam. Consequently, we extracted a working set of 2 million ham messages that were sent in the beginning of January, 2012 and still existed in the recipient's inbox. To reduce the possibility of collecting messages sent by uncaught spammers to inactive or dormant users, we selected messages only from users and to recipients who had logged in within the last two weeks. As in the case of spam, we collected two million messages with associated account data and merged them into text files for feature extraction.

@@

We then turned to LIBSVM, a library for SVM training and classification in C++ and Java, with wrappers for 
Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters 
($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$). We used Perl to convert the sparse feature 
matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package'
s tools to perform simple scaling on the data.

@@

@@

>>>>

Unlike spam messages, we did not have access to a comparable corpus of human-labeled ham 
messages. To simplify the labeling of ham messages, we made the assumption that messages remaining in 
the inboxes of active users after a period of several months would most likely not be spam. 
Consequently, we extracted a working set of 2 million ham messages that were sent in late December 2011 and early 
January 2012 and still existed in the recipient's inbox in March 2012. To reduce the 
possibility of collecting messages sent by uncaught spammers to inactive or dormant users, 
we selected messages only from users and to recipients who had logged in within the last two 
weeks. As in the case of spam, we collected two million messages with associated account 
data and merged them into text files for feature extraction.

@@

We then turned to LIBSVM, a library for SVM training and classification in C++ and Java, with wrappers for 
Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters 
($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$). We used Perl to convert the sparse feature 
matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package's 
tools to perform simple scaling on the data.

@@

Our initial results using the Pegasos SVM solver included in ScalaNLP were disappointing. The 
classifier trained on the bag-of-words feature set (from 22k messages) yielded a 77\% mean accuracy 
over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
51\%. Table \{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
feature models. We see that for the bag-of-words classifier, the words "year," "new," "happy,"  
"2012," "wish," and "you" all have high weights, which makes sense given that the ham data was 
collected around the beginning of January.

@@

\subsection{Data Collection}

The ham message collection methodology used in this project may have been flawed, leading questionable 
performance for several classifiers. The 2 million ham messages were sent over a span of 10 days, from 
December 23, 2011 to January 1, 2012. While we chose the corpus' start date as exactly four months before 
extracting the data, the coincidental timing with a global holiday seems to have introduced unusual 
skew in the bag-of-words features. In contrast, the majority of messages in the spam corpus span a period of 
7 months, with only a slight linear increase in volume over the interval.  Consequently, as seen in 
the Linear Regression and SVM feature weight tables, words typically associated with the New Year gained 
undue wieght during classification.  We plan to address this issue by creating a new ham message corpus 
through sampling that matches the temporal distribution of spam messages more closely.
